\textbf{ Discovery of a disrupting open cluster far into the Milky Way halo:\\ a recent star formation event in the leading arm of the Magellanic stream?}

### 2018-11-14 ms.tex
chopping

The halo of the Milky Way is typically characterized by its old ($\gtrsim 10~\textrm{Gyr}$), metal-poor ($\feh \approx -1.5$) stellar population.
This is understood as a signature of the dominant (in stellar mass) progenitor systems that were accreted early on in the formation of the Galaxy \citep[massive dwarf galaxies][]{Deason:2015, Fiorentino:2015}.
It is thought that these systems came in with significant gas reservoirs, but were quenched and stripped through collisional processes that heated and dispersed the gas \citep[e.g.,][]{Mayer:2006}, thus preventing immediate star formation in the deposited gas.
The Milky Way, however, continues to accrete satellite galaxies, as is evidenced by the prominent stellar stream from the Sagittarius dwarf galaxy \citep{Ibata:1994, Majewski:2003}, the presence of the Large and Small Magellanic Clouds (LMC and SMC), and about 50 dwarf satellites within the virial radius of the Galaxy.
While Sagittarius was likely stripped of its neutral gas long ago through repeated passage about the Milky Way \citep{Burton:1999, Tepper-Garcia:2018}, the \lmcsmc\ system is associated with $\approx 8\times 10^8~\msun$ of \hi gas \citep{Bruns:2005}, which extends into leading and trailing gas streams \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.
Here, we report the first discovery of a young cluster of co-moving stars in the Milky Way halo that appears to be associated with this \lmcsmc\ gas stream, which suggests that young stars can form from tidally stripped gas during low-mass mergers and may therefore exist throughout the otherwise aging stellar halo of our Galaxy.

The Magellanic stream (MS), including gas in the leading arm (LA), is a large stream of predominantly hydrogen gas emanating from the \lmcsmc\ system that contains a significant fraction of the total gas mass associated with the \lmcsmc\ \citep{Bruns:2005} and wraps nearly $\approx 200^\circ$ around the sky \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.
The trailing MS has been studied in great detail by large-area and high-resolution radio sky surveys: Recent surveys have found small-scale structure and gas fragmentation \citep[e.g.,][]{Nidever:2008, For:2014} and a large-scale bifurcation, with kinematically \citep{Nidever:2008} and chemically \citep{Fox:2013} distinct ``strands'' that lead back to the LMC and the SMC.
The LA gas has been found to connect to regions of low-column-density gas ($N\sim 10^{18}$--$10^{19}~\textrm{cm}^{-2}$) on the other side of the Galactic disk \citep{Putman:1998, Nidever:2008}, and has been decomposed into distinct gas features named LA I--IV \citep{Bruns:2005, Nidever:2008, Venzmer:2012}.
%that contain a small fraction of the total MS mass \citep[$\sim 4 \times10^7~\msun$;][]{Venzmer:2012}.

The origin and formation of the LA features of the MS is still uncertain.
Initial studies of the LA argued that the features closest to the \lmcsmc\ can be traced back to the SMC \citep{Putman:1998}, but it has also been argued that outer features of the LMC appear to lead directly into the LA I feature, implying an LMC origin \citep{Nidever:2008}.
Recent chemical abundance measurements along several sight-lines passing through the LA again suggested the SMC origin \citep{Fox:2018, Richter:2018}.
Whatever the origin of the gas, it is clear that tidal stripping by the Milky Way is required to form the LA \citep{Nidever:2008, Besla:2012}.
However, the LA features deviate from the predicted orbit of the \lmcsmc, implying that ram pressure or interactions with gas in the outer Milky Way disk (from the recent disk plane passage) may have removed orbital energy from the LA gas \citep[e.g.,][]{Bekki:2008}.

% Streams formed from tidally stripped material --- such as the Sagittarius or Palomar 5 \citep{Odenkirchen:2001} stellar streams, or gas streams such as the MS --- encode information about the orbital history and future trajectory of their progenitor system \citep[e.g.,][]{Johnston:1999}.
% For this reason, streams of debris are of interest for constraining the dark matter distribution around the Galaxy, as well as for providing a fossil record of the Milky Way's accretion history \citep{Johnston:1998, Bullock:2005}.
% Early models for the formation of the MS invoke ram pressure and tidal stripping to explain the sky distribution and extent of the MS, and typically require that the \lmcsmc\ system complete several orbits around the Galaxy in order to strip enough material \citep[e.g.,][]{Murai:1980, Gardiner:1994, Mastropietro:2005}.
% However, later
The gas in the MS therefore encodes information about the orbital history and future trajectory of the \lmcsmc, and about interactions between this gas and the Milky Way.
Combined with recent proper motion measurements of the LMC \citep{Kallivayalil:2006, Kallivayalil:2013} and improved models for the \lmcsmc\ that suggest they are on their first passage through the Galaxy \citep{Besla:2007, Besla:2010, Besla:2012}, several groups have used the MS to constrain properties of both the interaction history of the \lmcsmc\ and the dark matter halo of the Milky Way (see recent review by \citealt{DOnghia:2016}).
In general, these new models for the formation of the MS rely on past interactions between the LMC and SMC to preprocess the Magellanic gas distribution before infall and eventual stripping by the tidal field of the Milky Way \citep{Besla:2012, Diaz:2012};
More recent hydrodynamical simulations of the MCs showed that the repeated encounters even after the infall between the MCs strip gas not only from the SMC, but also from the LMC, and this tidally stripped gas can create filamentary gas structures both in the leading and trailing MS \citep{Pardy:2018}. Many observational studies \citep[e.g.,][]{Olsen:2011, Noel:2013, Mackey:2016, Carrera:2017, Choi:2018a, Choi:2018b, Zivick:2018, Belokurov:2019} have been made in the Magellanic system that are available to provide strong constraints on modeling the interaction history of the \lmcsmc\ in detail.

One critical difficulty in using the MS to further improve models of the \lmcsmc\ interaction and infall is the lack of distance information along the MS.
No significant over-density of stars have been found associated with the trailing MS \citep{Guhathakurta:1998}, thus leaving distance and tangential velocity information largely unknown.
While a small number of young OB stars have been found in the vicinity of the LA gas \citep{Casetti-Dinescu:2014, Zhang:2017}, given their sparsity and concentration near the Galactic plane, it is difficult to unambiguously associate them with the MS (they could instead be runaway OB stars from the Milky Way disk).

@@

% Notebook: Proper-motion-membership
\begin{figure*}[t!]
\centering
\includegraphics[width=0.9\textwidth]{figures/pm-model.pdf}
\caption{\textbf{Left panel}: Sky positions of all \gaia\ sources in this region with $\bprp < 0.35$ (grey markers).
The circles show the regions used as cluster and control fields in the proper motion inference (\sectionname~\ref{sec:pmclean}).
The control fields, together, have the same area as the cluster field.
\textbf{Right panel}: The colored 2D density shows the difference in number of sources per pixel in the cluster field versus the two control fields (see left panel).
The clear over-density in the cluster field (blue) are stars in \clustername. \yumi{can you add the mean proper motion vector of the cluster on this figure or figure 1?}}
\label{fig:pm-members}
\end{figure*}

@@

We measure the mean proper motion of \clustername\ by constructing a probabilistic model of the cluster and background populations using astrometric data from \gaia.
We start by selecting all stars with $(\bprp)_0 < 0.35$ to remove low-mass and old main sequence star contamination in the region.
\figurename~\ref{fig:pm-members}, left (grey points), shows the sky positions of stars that pass this blue cut in the region around the young cluster.
The larger solid-line circle indicates the region we define as the cluster area, and the two smaller dashed-line circles indicate control fields that are combined and used for modeling the background distribution of proper motions.
The control fields are designed to, together, have the same total area as the cluster field, and were chosen to have similar latitudes as the cluster field in the Magellanic stream coordinate system \citep{Nidever:2008}. \yumi{by assuming a smooth background stellar density at similar latitudes?}
The right panel of \figurename~\ref{fig:pm-members} shows the difference of the 2D proper motion distributions in the cluster and summed control fields.
The distinct over-density of stars near $(\mu_\alpha, \mu_\delta) \approx (-0.5, 0.5)~\masyr$ is the identified comoving association of blue stars.\footnote{Throughout this article, we use $\mu_\alpha$ to refer to the proper motion value provided by \gaia, which includes the $\cos\delta$ term.}
Note that \figurename~\ref{fig:pm-members} is only meant as an illustration: The cluster proper motion is determined probabilistically by taking into account the full \gaia\ covariance matrices for each source, as described below.

To measure the cluster (mean) proper motion, we first construct a model for the error-deconvolved proper motion distribution in the control fields using ``extreme deconvolution'' \citep[XD;][]{Bovy:2011} with two Gaussian components. XD takes into account the full error distributions for each proper motion measurement $\bs{\mu} = (\mu_\alpha, \mu_\delta)$, including covariances $\mat{C}_\mu$, provided by \gaia\ \DR{2}.
After running XD, we fix the parameters of the background model \yumi{you meant the model for the control fields? If so, please clarify that background model equals control fields model.} and model the proper motion distribution in the cluster region using a two-component mixture model with a single, isotropic Gaussian component for the error-deconvolved cluster distribution with mean $\bs{x}$ and isotropic variance $s^2$, and the XD-inferred model for the control field, $p_{\textrm{XD}}$, for the background component.
In detail, taking $f$ to be the fraction of blue stars in this region belonging to the young cluster, and \truepm\ to be the true proper motion for a single star,
\begin{align}
    p(\bs{\mu} \given \truepm, \mat{C}_\mu) &=
        \norm(\bs{\mu} \given \truepm, \mat{C}_\mu)\\
    p(\truepm \given f, \bs{x}, s) &=
        f \, p_{\textrm{cl}}(\truepm \given \bs{x}, s)
        + (1-f) \, p_{\textrm{XD}}(\truepm)\\
    p_{\textrm{cl}}(\truepm \given \bs{x}, s) &=
        \norm(\truepm \given \bs{x}, s^2 \, \mathbb{I})
\end{align}
where $\norm(\cdot \given \bs{y}, \mat{C})$ represents the multidimensional normal distribution with mean $\bs{y}$ and covariance matrix $\mat{C}$, and $\mathbb{I}$ is the identity matrix.
Because all distributions are Gaussian, the per-star parameters (the true proper motions, \truepm) can be analytically marginalized out so that the per-star likelihood can be expressed as $p(\bs{\mu} \given f, \bs{x}, s, \mat{C}_\mu)$.
We assume that the measurements for each star, $n$, are independent so that the full likelihood of all $N$ stars given the parameters $(f, \bs{x}, s)$ is
\begin{align}
    p(\{\bs{\mu}_n\}_N \given f, \bs{x}, s, \{\mat{C}_{\mu, n}\}_N) &=
        \prod_n^N p(\bs{\mu}_n \given f, \bs{x}, s, \mat{C}_{\mu, n}) \quad .
        \label{eq:likelihood}
\end{align}

We use an ensemble Markov chain Monte Carlo (MCMC) sampler \citep[\texttt{emcee};][]{emcee, Goodman:2010} to generate posterior samples over the parameters $(f, \bs{x}, s)$ using the likelihood defined above (\equationname~\ref{eq:likelihood}), and assuming the following prior probability distributions: uniform over the domain $(-5, 5)~\masyr$ for each component of $\bs{x}$, uniform in $f$, and uniform in log-$s$ over the domain $-6 < \ln s < 4$ (with $s$ in units of \masyr).
We run the sampler with 32 walkers for 256 steps as burn-in, then reset the sampler and run for an additional 512 steps. \yumi{why this number of walkers and steps? depends on number of fitting parameters and autocorrelation time?}
We then thin \yumi{thin? out of my own curiosity: why thinning is needed?} the resulting chains by taking every 16th sample, leaving a total of 1024 samples; We use these samples to estimate the median posterior parameter values and uncertainties,
For the young cluster, we find $\bs{x} = (-0.56,  0.47) \pm (0.04, 0.02)~\masyr$, $\ln s = -3.8 \pm 0.9$, and $f = 0.14 \pm 0.02$.
The proper-motion dispersion is consistent with zero, and is likely unmeasured, because of the proper motion uncertainties; 95\% of the posterior samples have $s < 0.09~\masyr$.
\tablename~\ref{tbl:gaiadata} contains a subset of the \gaia\ \DR{2} data for stars in the cluster field, with kinematic membership probabilities determined from this modeling procedure.

@@

In detail, we start by using the \texttt{isochrones} package \citep{Morton:2015} to generate posterior samplings over stellar parameters for each individual source given its photometry and an interim prior.
We use the \acronym{MIST} \citep{Dotter:2016, Choi:2016, Paxton:2011, Paxton:2013, Paxton:2015} isochrone grid, and \texttt{isochrones} automatically performs interpolation between the provided grid of stellar isochrones to predict photometry given a set of stellar parameters.
Here, the stellar parameters for each source are its ``equal evolutionary point'' number $\eep$ (see \acronym{MIST} documentation\footnote{\url{http://waps.cfa.harvard.edu/MIST/}}), age $\tau$, metallicity $\feh$, extinction $A_V$, and distance $D$, which can be uniquely mapped to a point in the observed CMD;
The likelihood of a given set of these parameters, $\bs{\theta} = (\eep, \tau, \feh, A_V, D)$, is then computed from the photometry and (assumed Gaussian) photometric uncertainties given the predicted photometry. \yumi{is it easy to use real photometric uncertainties instead of assuming Gaussian? If not, we can wait what the referee will say.}

To generate posterior samples, we must also specify prior probability distributions over the parameters $\bs{\theta}$.
The priors for each parameter are summarized in \tablename~\ref{tbl:priors}.
The bounds on $\eep$ limit the isochrone to evolutionary phases between the zero-age main sequence to the terminal-age main sequence, but this prior is actually computed using the stellar mass computed from the isochrone parameters (see caption of \tablename~\ref{tbl:priors}).
The prior and bounds on $A_V$ are set to prefer small and reasonable values of extinction for this moderately high Galactic latitude region. \yumi{Av is known to follow a log-normal distribution.}
The prior on distance assumes a uniform space density of stars.

@@

We find that the photometry for individual sources in the lower main sequence are very poorly constrained in all parameters, and the prior tends to pull the posterior samplings to prefer closer, older stellar parameters.
This is a weakness of our methodology for performing the hierarchical inference: Each source is considered independently, even though there is clear structure in the CMD (i.e. \figurename~\ref{fig:decam-cmd}), and inferring isochronal parameters for individual lower main sequence stars is a fundamentally degenerate problem.
A more correct way to do this would be to infer the stellar parameters of \emph{all} stars, the cluster hyperparameters, and the background simultaneously.
However, for the \Nisofit\ stars we are using (in the red box in \figurename~\ref{fig:decam-cmd}), this model would have $\sim$2,000 free parameters if left unmarginalized.
We are developing tools to perform star cluster parameter inference in this way (Morton et al., in prep.), but here we adopt a simple hack to allow us to instead perform individual posterior samplings and then combine those samplings into a hierarchical inference. \yumi{I like this fancy way. But it would be helpful for readers (potential users of this method) if you can briefly describe the strength of this methodology over the standard isochrone fitting for a star cluster?}

The brightest stars in the cluster have very precise \decam\ photometry, and have precise astrometry from \gaia.
Given their location in the CMD, these stars must be young, and so also have much better constrained stellar parameters. \yumi{I am not convinced by that bright stars have better constrained stellar parameters. Their color and magnitude are sensitive to dust and effective temperature, but their age and metallicity would not constrained better than fainter stars. I suggest to remove the second half of the previous sentence.}
We therefore use the brightest \emph{kinematic} member of \clustername\ as an ``anchor'' star:
We model every other source in the \decam\ selection region (i.e. excluding this one) by fitting the photometry of it and the anchor star simultaneously, assuming they have the same age, metallicity, distance, and extinction but different \eep\ values.
Motivated by the possible signature of an (unresolved) binary sequence in the \decam\ CMD, we add one further piece of complexity to the model by allowing the photometry of each non-anchor source to be fit as an unresolved binary star system.
This adds an additional parameter, the unresolved binary mass ratio $q$, to the list of inferred stellar parameters for each source, but, for this work, we ignore the binary companions and implicitly marginalize over $q$ in what follows.
The details of this model are handled by the \texttt{isochrones} package.

@@

We use the posterior samples from the hierarchical inference described in \sectionname~\ref{sec:popmodel} to compute the posterior probabilities that each source in the \decam\ CMD selection box (red box, \figurename~\ref{fig:decam-cmd}) is a member of the cluster.
\figurename~\ref{fig:hierarch-iso} shows the \decam\ photometry for all sources with membership probability $> 0.5$ (circle markers).
Over-plotted in \figurename~\ref{fig:hierarch-iso} (blue, solid line) is a \acronym{MIST} isochrone with the median posterior parameters derived from the hierarchical modeling, shifted to the median distance and extincted given the median $A_V$ value. \yumi{How our Av is different from Schlegel map (https://irsa.ipac.caltech.edu/cgi-bin/bgTools/nph-bgExec)? Their Av is about 0.2. Are we consistent with them within uncertainty?}
We find that \clustername\ is indeed young, distant, and metal poor, with median posterior values of age $\tau \approx \clage$, distance $D \approx \cldist$, and metallicity $\feh \approx \clfeh$ (see also \tablename~\ref{tbl:clusterparams} for a summary).
Also plotted in \figurename~\ref{fig:hierarch-iso} is the equal-mass binary sequence (green, dashed line) computed from the median posterior sample: The abundance of sources between the nominal isochrone and the binary sequence highlights the fact that the cluster may contain a significant number of binary or multiple star systems, but we leave a detailed study of multiplicity to future work.

@@

The discovery of \clustername\ provides a critical distance constraint to the LA and will aid future Magellanic system and Milky Way modeling efforts.
It also provides an opportunity to study star formation in a unique environment (i.e., low gas density and low metallicity), unlike that of the Milky Way disk or any other cluster-forming region.
The serendipitous discovery of this cluster is a reminder that the combined value of the \gaia\ data with deep, large-area imaging surveys provides a wealth of information about our Galaxy and stellar halo.

>>>>

The halo of the Milky Way is typically characterized by its old ($\gtrsim 10~\textrm{Gyr}$), metal-poor ($\feh \approx -1.5$) stellar population.
This is understood as a signature of the dominant (in stellar mass) progenitor systems that were accreted early on in the formation of the Galaxy \citep[massive dwarf galaxies][]{Deason:2015, Fiorentino:2015}.
It is thought that these systems came in with significant gas reservoirs, but were quenched and stripped through collisional processes that heated and dispersed the gas \citep[e.g.,][]{Mayer:2006}, thus preventing immediate star formation in the deposited gas.
The Milky Way, however, continues to accrete satellite galaxies, as is evidenced by the prominent stellar stream from the Sagittarius dwarf galaxy \citep{Ibata:1994, Majewski:2003}, the presence of the Large and Small Magellanic Clouds (LMC and SMC), and about 50 dwarf satellites within the Galactic halo.
While Sagittarius was likely stripped of its neutral gas long ago \citep{Burton:1999, Tepper-Garcia:2018}, the \lmcsmc\ system is associated with $\approx 8\times 10^8~\msun$ of \hi gas \citep{Bruns:2005}, which extends into leading and trailing gas streams \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.
Here we report the first discovery of a young star cluster in the Milky Way halo that appears to be associated with this \lmcsmc\ gas stream, which suggests that young stars can form from tidally stripped gas during low-mass mergers and may therefore exist throughout the otherwise aging stellar halo.

The Magellanic stream (MS), including gas in the leading arm (LA), is a large stream of predominantly hydrogen gas emanating from the \lmcsmc\ system that wraps nearly $\approx 200^\circ$ around the sky \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010} and contains a significant fraction of the total gas mass associated with the \lmcsmc\ \citep{Bruns:2005}.
The trailing MS has been studied in great detail by large-area and high-resolution radio sky surveys: Recent surveys have found small-scale structure and gas fragmentation \citep[e.g.,][]{Nidever:2008, For:2014} and a large-scale bifurcation, with kinematically \citep{Nidever:2008} and chemically \citep{Fox:2013} distinct ``strands'' that lead back to the LMC and the SMC.
The LA gas has been found to connect to regions of low-column-density gas ($N\sim 10^{18}$--$10^{19}~\textrm{cm}^{-2}$) on the other side of the Galactic disk \citep{Putman:1998, Nidever:2008}, and has been decomposed into distinct gas features named LA I--IV \citep{Bruns:2005, Nidever:2008, Venzmer:2012}.
%that contain a small fraction of the total MS mass \citep[$\sim 4 \times10^7~\msun$;][]{Venzmer:2012}.

The origin and formation of these LA features is still uncertain.
Initial studies of the LA argued that the features closest to the \lmcsmc\ can be traced back to the SMC \citep{Putman:1998}, but outer features of the LMC appear to lead directly into the LA I feature \citep{Nidever:2008}.
Recent chemical abundance measurements along several sight-lines passing through the LA again support an SMC origin for the gas \citep{Fox:2018, Richter:2018}.
Whatever the origin of the LA gas, it is clear that tidal stripping by the Milky Way is required to form the LA \citep{Nidever:2008, Besla:2012}.
However, the LA features deviate from the predicted orbit of the \lmcsmc, implying that ram pressure or interactions the outer Milky Way disk (from the recent disk plane passage) may have removed orbital energy from the LA gas \citep[e.g.,][]{Bekki:2008}.

% Streams formed from tidally stripped material --- such as the Sagittarius or Palomar 5 \citep{Odenkirchen:2001} stellar streams, or gas streams such as the MS --- encode information about the orbital history and future trajectory of their progenitor system \citep[e.g.,][]{Johnston:1999}.
% For this reason, streams of debris are of interest for constraining the dark matter distribution around the Galaxy, as well as for providing a fossil record of the Milky Way's accretion history \citep{Johnston:1998, Bullock:2005}.
% Early models for the formation of the MS invoke ram pressure and tidal stripping to explain the sky distribution and extent of the MS, and typically require that the \lmcsmc\ system complete several orbits around the Galaxy in order to strip enough material \citep[e.g.,][]{Murai:1980, Gardiner:1994, Mastropietro:2005}.
% However, later
The gas in the MS encodes information about the past and future trajectory of the \lmcsmc, and about interactions between this gas and the Milky Way.
Combined with recent proper motion measurements of the LMC \citep{Kallivayalil:2006, Kallivayalil:2013} and improved models for the \lmcsmc\ that suggest they are on their first passage through the Galaxy \citep{Besla:2007, Besla:2010, Besla:2012}, several groups have used the MS to constrain properties of both the interaction history of the \lmcsmc\ and of the dark matter halo of the Milky Way (see recent review by \citealt{DOnghia:2016}).
In general, these models for the formation of the MS rely on past interactions between the LMC and SMC to preprocess the Magellanic gas distribution before infall and eventual stripping by the tidal field of the Milky Way \citep{Besla:2012, Diaz:2012};
More recent hydrodynamical simulations of the MCs showed that the repeated encounters between the \lmcsmc\ strip gas both from the SMC and the LMC, and this tidally stripped gas can then create filamentary structures both in the leading and trailing MS \citep{Pardy:2018}.
Many observational studies \citep[e.g.,][]{Olsen:2011, Noel:2013, Mackey:2016, Carrera:2017, Choi:2018a, Choi:2018b, Zivick:2018, Belokurov:2019} of the \lmcsmc\ themselves will provide strong constraints the interaction history of the MCs.

One critical difficulty in using the MS to further improve models of the \lmcsmc\ interaction and infall is the lack of distance information along the MS.
No significant over-density of stars have been found associated with the trailing MS \citep{Guhathakurta:1998}, thus leaving distance and tangential velocity information unknown.
While a small number of OB stars have been found in the vicinity of the LA gas \citep{Casetti-Dinescu:2014, Zhang:2017}, their sparsity and concentration near the Galactic plane make it difficult to unambiguously associate them with the MS and not runaway OB stars from the Milky Way disk.

@@

% Notebook: Proper-motion-membership
\begin{figure*}[t!]
\centering
\includegraphics[width=0.9\textwidth]{figures/pm-model.pdf}
\caption{\textbf{Left panel}: Sky positions of all \gaia\ sources in this region with $\bprp < 0.35$ (grey markers).
The circles show the regions used as cluster and control fields in the proper motion inference (\sectionname~\ref{sec:pmclean}).
The control fields, together, have the same area as the cluster field.
\textbf{Right panel}: The colored 2D density shows the difference in number of sources per pixel in the cluster field versus the two control fields (see left panel).
The clear over-density in the cluster field (blue) are stars in \clustername.}
\label{fig:pm-members}
\end{figure*}

@@

We measure the mean proper motion of \clustername\ by constructing a probabilistic model of the cluster and background populations using astrometric data from \gaia.
We start by selecting all stars with $(\bprp)_0 < 0.35$ to remove low-mass and old main sequence star contamination in the region.
\figurename~\ref{fig:pm-members}, left (grey points), shows the sky positions of stars that pass this blue cut in the region around the young cluster.
The larger solid-line circle indicates the region we define as the cluster area, and the two smaller dashed-line circles indicate control fields that are combined and used for modeling the background distribution of proper motions.
The control fields are designed to, together, have the same total area as the cluster field, and were chosen to have similar latitudes as the cluster field in the Magellanic stream coordinate system \citep{Nidever:2008}.
We assume that the background density in the cluster field is equivalent to the average background density of the join control fields.
From visual inspection of the CMDs of the control fields, we do not see any significant clustered over-density and therefore assume that these fields are dominated by the background stellar density.
The right panel of \figurename~\ref{fig:pm-members} shows the difference of the 2D proper motion distributions in the cluster and summed control fields.
The distinct over-density of stars near $(\mu_\alpha, \mu_\delta) \approx (-0.5, 0.5)~\masyr$ is the identified comoving association of blue stars.\footnote{Throughout this article, we use $\mu_\alpha$ to refer to the proper motion value provided by \gaia, which includes the $\cos\delta$ term.}
Note that \figurename~\ref{fig:pm-members} is only meant as an illustration: The cluster proper motion is determined probabilistically by taking into account the full \gaia\ covariance matrices for each source, as described below.

To measure the cluster (mean) proper motion, we first construct a model for the error-deconvolved proper motion distribution in the control fields using ``extreme deconvolution'' \citep[XD;][]{Bovy:2011} with two Gaussian components. XD takes into account the full error distributions for each proper motion measurement $\bs{\mu} = (\mu_\alpha, \mu_\delta)$, including covariances $\mat{C}_\mu$, provided by \gaia\ \DR{2}.
After running XD on the proper motion distribution of the control fields, we fix the parameters of the density model and use this as the background model for the cluster field.
We model the proper motion distribution in the cluster region using a two-component mixture model with a single, isotropic Gaussian component for the error-deconvolved cluster distribution with mean $\bs{x}$ and isotropic variance $s^2$, and the XD-inferred background model, $p_{\textrm{XD}}$, for the background component.
In detail, taking $f$ to be the fraction of blue stars in this region belonging to the young cluster, and \truepm\ to be the true proper motion for a single star,
\begin{align}
    p(\bs{\mu} \given \truepm, \mat{C}_\mu) &=
        \norm(\bs{\mu} \given \truepm, \mat{C}_\mu)\\
    p(\truepm \given f, \bs{x}, s) &=
        f \, p_{\textrm{cl}}(\truepm \given \bs{x}, s)
        + (1-f) \, p_{\textrm{XD}}(\truepm)\\
    p_{\textrm{cl}}(\truepm \given \bs{x}, s) &=
        \norm(\truepm \given \bs{x}, s^2 \, \mathbb{I})
\end{align}
where $\norm(\cdot \given \bs{y}, \mat{C})$ represents the multidimensional normal distribution with mean $\bs{y}$ and covariance matrix $\mat{C}$, and $\mathbb{I}$ is the identity matrix.
Because all distributions are Gaussian, the per-star parameters (the true proper motions, \truepm) can be analytically marginalized out so that the per-star likelihood can be expressed as $p(\bs{\mu} \given f, \bs{x}, s, \mat{C}_\mu)$.
We assume that the measurements for each star, $n$, are independent so that the full likelihood of all $N$ stars given the parameters $(f, \bs{x}, s)$ is
\begin{align}
    p(\{\bs{\mu}_n\}_N \given f, \bs{x}, s, \{\mat{C}_{\mu, n}\}_N) &=
        \prod_n^N p(\bs{\mu}_n \given f, \bs{x}, s, \mat{C}_{\mu, n}) \quad .
        \label{eq:likelihood}
\end{align}

We use an ensemble Markov chain Monte Carlo (MCMC) sampler \citep[\texttt{emcee};][]{emcee, Goodman:2010} to generate posterior samples over the parameters $(f, \bs{x}, s)$ using the likelihood defined above (\equationname~\ref{eq:likelihood}), and assuming the following prior probability distributions: uniform over the domain $(-5, 5)~\masyr$ for each component of $\bs{x}$, uniform in $f$, and uniform in log-$s$ over the domain $-6 < \ln s < 4$ (with $s$ in units of \masyr).
We run the sampler with 32 walkers for 256 steps as burn-in, then reset the sampler and run for an additional 512 steps, after which the chains appear converged.
We then downsample the resulting chains by taking every 16th sample to preserve closer-to-independent samples, leaving a total of 1024 samples; We use these samples to estimate the median posterior parameter values and uncertainties,
For the young cluster, we find $\bs{x} = (-0.56,  0.47) \pm (0.04, 0.02)~\masyr$, $\ln s = -3.8 \pm 0.9$, and $f = 0.14 \pm 0.02$.
The proper-motion dispersion is consistent with zero, and is likely unmeasured, because of the proper motion uncertainties; 95\% of the posterior samples have $s < 0.09~\masyr$.
\tablename~\ref{tbl:gaiadata} contains a subset of the \gaia\ \DR{2} data for stars in the cluster field, with kinematic membership probabilities determined from this modeling procedure.

@@

In detail, we start by using the \texttt{isochrones} package \citep{Morton:2015} to generate posterior samplings over stellar parameters for each individual source given its photometry and an interim prior.
We use the \acronym{MIST} \citep{Dotter:2016, Choi:2016, Paxton:2011, Paxton:2013, Paxton:2015} isochrone grid, and \texttt{isochrones} automatically performs interpolation between the provided grid of stellar isochrones to predict photometry given a set of stellar parameters.
Here, the stellar parameters for each source are its ``equal evolutionary point'' number $\eep$ (see \acronym{MIST} documentation\footnote{\url{http://waps.cfa.harvard.edu/MIST/}}), age $\tau$, metallicity $\feh$, extinction $A_V$, and distance $D$, which can be uniquely mapped to a point in the observed CMD;
The likelihood of a given set of these parameters, $\bs{\theta} = (\eep, \tau, \feh, A_V, D)$, is then computed from the photometry and (assumed Gaussian) photometric uncertainties given the predicted photometry. % \yumi{is it easy to use real photometric uncertainties instead of assuming Gaussian? If not, we can wait what the referee will say.}
% Not easy, no - these are what David provides from his pipeline, and what Isochrones assumes as input. But the photometry is so precise the errors really are close to Gaussian

To generate posterior samples, we must also specify prior probability distributions over the parameters $\bs{\theta}$.
The priors for each parameter are summarized in \tablename~\ref{tbl:priors}.
The bounds on $\eep$ limit the isochrone to evolutionary phases between the zero-age main sequence to the terminal-age main sequence, but this prior is actually computed using the stellar mass computed from the isochrone parameters (see caption of \tablename~\ref{tbl:priors}).
The prior and bounds on $A_V$ are set to prefer small and reasonable values of extinction for this moderately high Galactic latitude region.
% \yumi{Av is known to follow a log-normal distribution.} Yea the referee could push back, but we'll see!
The prior on distance assumes a uniform space density of stars.

@@

We find that the photometry for individual sources in the lower main sequence are very poorly constrained in all parameters, and the prior tends to pull the posterior samplings to prefer closer, older stellar parameters.
This is a weakness of our methodology for performing the hierarchical inference: Each source is considered independently, even though there is clear structure in the CMD (i.e. \figurename~\ref{fig:decam-cmd}), and inferring isochronal parameters for individual lower main sequence stars is a fundamentally degenerate problem.
A more correct way to do this would be to infer the stellar parameters of \emph{all} stars, the cluster hyperparameters, and the background simultaneously.
However, for the \Nisofit\ stars we are using (in the red box in \figurename~\ref{fig:decam-cmd}), this model would have $\sim$2,000 free parameters if left unmarginalized.
We are developing tools to perform star cluster parameter inference in this way (Morton et al., in prep.), but here we adopt a simple hack to allow us to instead perform individual posterior samplings and then combine those samplings into a hierarchical inference.
% \yumi{I like this fancy way. But it would be helpful for readers (potential users of this method) if you can briefly describe the strength of this methodology over the standard isochrone fitting for a star cluster?}
% What is standard isochrone fitting?

The brightest stars in the cluster have very precise \decam\ photometry, and have precise astrometry from \gaia.
Given their location in the CMD, these stars must be young.
We therefore use the brightest \emph{kinematic} member of \clustername\ as an ``anchor'' star:
We model every other source in the \decam\ selection region (i.e. excluding this one) by fitting the photometry of it and the anchor star simultaneously, assuming they have the same age, metallicity, distance, and extinction but different \eep\ values.
Motivated by the possible signature of an (unresolved) binary sequence in the \decam\ CMD, we add one further piece of complexity to the model by allowing the photometry of each non-anchor source to be fit as an unresolved binary star system.
This adds an additional parameter, the unresolved binary mass ratio $q$, to the list of inferred stellar parameters for each source, but, for this work, we ignore the binary companions and implicitly marginalize over $q$ in what follows.
The details of this model are handled by the \texttt{isochrones} package.

@@

We use the posterior samples from the hierarchical inference described in \sectionname~\ref{sec:popmodel} to compute the posterior probabilities that each source in the \decam\ CMD selection box (red box, \figurename~\ref{fig:decam-cmd}) is a member of the cluster.
\figurename~\ref{fig:hierarch-iso} shows the \decam\ photometry for all sources with membership probability $> 0.5$ (circle markers).
Over-plotted in \figurename~\ref{fig:hierarch-iso} (blue, solid line) is a \acronym{MIST} isochrone with the median posterior parameters derived from the hierarchical modeling, shifted to the median distance and extincted given the median $A_V$ value.
We find that \clustername\ is indeed young, distant, and metal poor, with median posterior values of age $\tau \approx \clage$, distance $D \approx \cldist$, and metallicity $\feh \approx \clfeh$ (see also \tablename~\ref{tbl:clusterparams} for a summary).
We find that the inferred extinction, $A_V \approx 0.33$, is somewhat larger than the value $A_{V, {\rm SFD}} \approx 0.2$ from the SFD dust map \citep{Schlegel:1998}.

Also plotted in \figurename~\ref{fig:hierarch-iso} is the equal-mass binary sequence (green, dashed line) computed from the median posterior sample: The abundance of sources between the nominal isochrone and the binary sequence highlights the fact that the cluster may contain a significant number of binary or multiple star systems, but we leave a detailed study of multiplicity to future work.

@@

The discovery of \clustername\ provides a critical distance constraint to the LA and will aid future Magellanic system and Milky Way modeling efforts.
It also provides an opportunity to study star formation in a unique environment (i.e. low gas density, low metallicity, and large velocity), unlike that of the Milky Way disk or any other local cluster-forming region.
The serendipitous discovery of this cluster is a reminder that the combined value of the \gaia\ data with deep, large-area imaging surveys provides a wealth of information about our Galaxy and stellar halo.
